MS_data_clean

### 2022-01-26 MS_data_modeling/MS_data_explore.Rmd
Further work on paper

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lifecycle)
library(rvest)
library(dplyr)
library(tidyr)
library(plotly)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(kableExtra)
library(rcol)
library(plyr)

MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_10-26-21_ALR seltmann-colab.xlsx', sheet = 2)

@@

@@

MS_data_format %>% filter(!(is.na(specializationMS)), is.na(pollenHosts)) ### there
```

@@

## pulls all unique maximum level taxonomic names from MS pollen hosts
ms_pollen_hosts <- ms_pollen_hosts[!(ms_pollen_hosts == '')]

ms_pollen_hosts
```

@@

## Make the below two entries have 0 rows with Avery's help
```{r}
family_key %>% filter(is.na(pollen_host_familys))
```

@@

@@

compare_df
```

@@

data.frame(data_sources, counts) %>% mutate('Percent of Data' = counts / nrow(compare_df)) %>% 
  dplyr::rename('Data Source' = data_sources, 'Unique Species Count' = counts) %>% kbl(caption = "Bee Data Sources Summary") %>%
  kable_styling()
```

@@

odd_generalists <- compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsFamilysMS, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 == 1)

@@

@@

odd_specialists <- compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsMS, pollenHostsFamilysMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 > 3)

odd_specialists

@@

@@

```{r}
modeling_df <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% mutate(specialist = ifelse(specializationMS == 'Polylectic', 1, 0 )) %>% dplyr::select(interestBee, nBeeGLOBEE, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, FamilyDegreeMS, specialist)

modeling_df
```

@@

```{r}
glmfit <- glm(as.factor(specialist) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl, family = 'binomial')
predicttrain_log = predict(glmfit, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="response")
predicttest_log = predict(glmfit, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="response")

@@

1 - train_err
1 - test_err
```
# Try multi feature with best model:

```{r}
rf_fit <-  randomForest(as.factor(specialist) ~ ., data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(-specialist), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(-specialist), type="class")

@@

# Begin Baysian Approaches

```{r}
modeling_df
```

```{r}
compare_df %>% filter(interestBee == 'Agapostemon Angelicus') 
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lifecycle)
library(rvest)
library(dplyr)
library(tidyr)
library(plotly)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(kableExtra)
library(rcol)
library(plyr)
library(caret)

MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/UCSB Archive/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 2)

@@

```{r}
interactions <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv')

(interactions %>% filter(targetTaxonRank == 'family') %>% dplyr::select(sourceTaxonName, interactionTypeName, targetTaxonName) %>% 
    dplyr::rename('Bee Species' = sourceTaxonName, 'Interaction' = interactionTypeName, 'Plant Family' = targetTaxonName))[c(172, 112, 300, 124 ,5),]  %>%
  kbl() %>% kable_styling()

```

@@

MS_data_format <- MS_data_format %>% filter(!(is.na(specializationMS)), !is.na(pollenHosts)) ### there

MS_data_format
```

@@

## pulls all unique maximum level taxonomic names from MS pollen hosts
ms_pollen_hosts <- ms_pollen_hosts[!(ms_pollen_hosts == '')]
```

@@

## Make the below two entries have 0 rows with Avery's help
```{r}
family_key %>% filter(is.na(pollen_host_familys) | (pollen_host_familys == 'NA'))
```

@@

### Note, NAs are a result of there being no pollen hsots listed for a bee (should be fixed by filtering out bees with no pollen hosts), and Unknowns are a results of bees not being able to be found in catalougue of life

@@

compare_df %>% filter(existsALL == 1) %>% dplyr::select(interestBee, specializationMS, FamilyDegreeMS, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, pollenHostsFamilysMS, pollenHostGLOBEE) %>%
  dplyr::rename('Bee Species' = interestBee, 'Binary Label MS' = specializationMS, 'Degree MS' = FamilyDegreeMS, 'Raw Degree GLOBI' = degreeGLOBEE, 
                'Degree Cut GLOBI' = degreeGLOBEE_C5, 'Tranformed Degree GLOBI' = degreeGLOBEE_T, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts GLOBI' = pollenHostGLOBEE) %>%
  kbl() %>% kable_styling()
```

@@

data.frame(data_sources, counts) %>% mutate('Percent of All Bee Species Documented' = counts / nrow(compare_df)) %>% 
  dplyr::rename('Data Source' = data_sources, 'Unique Species Count' = counts) %>% kbl(caption = "Bee Data Sources Summary") %>%
  kable_styling()
```

@@

odd_generalists <- compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5,degreeGLOBEE_T, degreeGLOBEE, pollenHostsFamilysMS, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 == 1)

@@

```{r}

(odd_generalists %>% dplyr:: select(interestBee, specializationMS, degreeGLOBEE_C5, degreeGLOBEE_T, pollenHostsFamilysMS, pollenHostGLOBEE) %>%
  dplyr::rename('Bee Species' = interestBee, 'Binary Label MS' = specializationMS, 
                'Degree Cut GLOBI' = degreeGLOBEE_C5, 'Tranformed Degree GLOBI' = degreeGLOBEE_T, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts GLOBI' = pollenHostGLOBEE))[c(3,5),] %>% 
  kbl(caption = 'Odd Generalists') %>% kable_styling()

```

@@

odd_specialists <- compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, degreeGLOBEE_T, pollenHostsMS, pollenHostsFamilysMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 > 3)

@@

```{r}
(odd_specialists %>% dplyr:: select(interestBee, specializationMS, degreeGLOBEE_C5, degreeGLOBEE_T, pollenHostsFamilysMS, pollenHostGLOBEE) %>%
  dplyr::rename('Bee Species' = interestBee, 'Binary Label MS' = specializationMS, 
                'Degree Cut GLOBI' = degreeGLOBEE_C5, 'Tranformed Degree GLOBI' = degreeGLOBEE_T, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts GLOBI' = pollenHostGLOBEE) )%>% 
  kbl(caption = 'Odd Specialists') %>% kable_styling()  
```

@@

```{r}
modeling_df <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% mutate(specialist = ifelse(specializationMS == 'Polylectic', 1, 0 )) %>% dplyr::select(interestBee, nBeeGLOBEE, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, FamilyDegreeMS, specialist, pollenHostsFamilysMS, pollenHostGLOBEE)

modeling_df
ggplot(data= modeling_df, aes(x = nBeeGLOBEE, y = degreeGLOBEE_T, color = as.factor(specialist))) + geom_point() + xlim(0, 1000)
```
```{r}
length(unique(word(modeling_df$interestBee, 1)))

pollen_hosts_represented <- c(modeling_df$pollenHostsFamilysMS, modeling_df$pollenHostGLOBEE)
length(unique(pollen_hosts_represented))
```

```{r}
set.seed(123)
```

@@

```{r}
glmfit <- cv.glmnet(modeling_df %>% dplyr::select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), as.factor(modeling_df$specialist), family = 'binomial', alpha=1, 
          nfolds = 10)

coef  <-  coef(glmfit) ## uses lowest mse lambda
coef

predicttrain_log = predict(glmfit, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")
predicttest_log = predict(glmfit, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")

@@

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")

@@

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")

# Try multi feature with best model:

```{r}
rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + degreeGLOBEE_T + degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")

1 - train_err
1 - test_err

misclassified <- tst.cl[tst.cl$specialist == predict_test,]
misclassified_vec <- tst.cl$specialist != predict_test

paste('Percent of misclassified that were specialists:', round(sum(misclassified$specialist) / nrow(misclassified), 3))
paste('Percent of studied that were specialists:', round(sum(modeling_df$specialist) / nrow(modeling_df), 3))
```

```{r}
tst.cl$misclass <- as.numeric(misclassified_vec)
splits <- c(0, 5, 10, 100, 100000)
tst.cl$nBeeBin <- as.factor(cut(tst.cl$nBeeGLOBEE, breaks=splits))
tst.cl %>% dplyr::select(misclass, nBeeBin) %>% group_by(nBeeBin) %>% dplyr::summarize(freq = n(), misclass_rate = mean(misclass))
```

# Try fitting to family degree:

```{r}
ggplot(data = modeling_df, aes(x= FamilyDegreeMS)) + geom_histogram()

splits <- c(0, 1, 2, 5, 10, 100)
modeling_df$MSFamilyDegreeBins <- cut(modeling_df$FamilyDegreeMS, breaks=splits)
modeling_df
```

tree <- tree(as.factor(MSFamilyDegreeBins) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl)
cv <- cv.tree(tree, rand = folds, FUN = prune.misclass, K = 10)

train_pred_tree <- predict(tree, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')
test_pred_tree <- predict(tree, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')

```{r}
glmfit <- cv.glmnet(modeling_df %>% dplyr::select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), as.factor(modeling_df$specialist), family = 'binomial', alpha=1, 
          nfolds = 10)

coef  <-  coef(glmfit) ## uses lowest mse lambda
coef

predicttrain_log = predict(glmfit, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")
predicttest_log = predict(glmfit, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")

# Try multi feature with best model:

```{r}
rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + degreeGLOBEE_T + degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")

```{r}
max(modeling_df$degreeGLOBEE_C5)
```

## Step 1: Classify bees as weird by finding misclassified bees by ML algoritem
## Step 2: Pull family level from both data and literature
## Step 3: Filter differences in family citations
## Step 4: Find similarities of family discrepencies (this identifies bias in data)
## Step 5:
